---
title: "Inference and Simulation"
author: "Yuxiao Li"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
---

# Overview
This file contains the code for inference and reproducing rainfall data. Most of results shown in the paper can be found here.

In general, we want to

* load the results from estimation
* simulate one replicate of rainfall base on skew-t model and Gaussian model
* visualize the simulations
* simulate multiple replicates of rainfall base on skew-t model and Gaussian model
* calculate the some important quantities related to rainfall
* visualize the results

# Librares and Data

## Libraries
```{r}
library(here)
library(geoR)
library(sn)
library(cowplot)
```

## load data
```{r load estimate values}
load(here("results", "estimation.cov.VAR1.RData"))
load(here("results", "rain.long.RData"))
raindata <- matrix(rain.long$rain[1:230400],28800,8)
#raindata <- matrix(rain.1min.df$rain[1:115200],14400,8)
NT <- nrow(raindata)
NS <- ncol(raindata)
sigma21.st <- results.VAR1$sigma21
beta1.st <- results.VAR1$beta1
dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
for (i in 1:NS) {
    for (j in 1:NS) {
        dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
    }
}
B1.st <-  sigma21.st * matern(dist,beta1.st,1)
alpha <- results.VAR1$alpha
b0.st <- results.VAR1$b0
b1.st <- results.VAR1$b1
ut <- results.VAR1$ut
```

```{r load estimate values norm}
load(here("results", "estimation.cov.VAR1.norm.RData"))
load(here("results", "rain.long.RData"))
raindata <- matrix(rain.long$rain[1:230400],28800,8)
#raindata <- matrix(rain.1min.df$rain[1:115200],14400,8)
NT <- nrow(raindata)
NS <- ncol(raindata)
sigma21.norm <- results.VAR1.norm$sigma21
beta1.norm <- results.VAR1.norm$beta1
dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
for (i in 1:NS) {
    for (j in 1:NS) {
        dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
    }
}
B1.norm <-  sigma21.norm * matern(dist,beta1.norm,1)
b0.norm <- results.VAR1.norm$b0
b1.norm <- results.VAR1.norm$b1
```


#Simulate one replicate

##Simulate one replicate from skew-t model
```{r simulate 1 error st}
error.st <- matrix(0,NT,NS)
set.seed(4)
for(s in 1 :NS){
    nu <- 4
    bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
        alpha/sqrt(1 + alpha^2)
    omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
    xi<- -omega * bvdelta
    error.st[,s] <- rst(NT, xi = xi, omega = omega, alpha, nu)
}
```

```{r simulate 1 rainfall st}
y.st <- matrix(0,NT,NS)
for(t in 2: NT){
    y.st[t,] <- B1.st %*% raindata[t - 1,] + error.st[t,] * (b0.st + b1.st * mean(raindata[t - 1,]))
    y.st[t,] <- ifelse(y.st[t,] > ut[t,], y.st[t,], 0)
}
```

##Simulate one replicate from Gaussian model

```{r simulate 1 error norm}
error.norm <- matrix(0,NT,NS)
set.seed(4)
for(s in 1 :NS) error.norm[,s] <- rnorm(NT)
```

```{r simulate 1 rainfall norm}
y.norm <- matrix(0,NT,NS)
for(t in 2: NT){
    y.norm[t,] <- B1.norm %*% raindata[t - 1,] + 
        error.norm[t,] * (b0.norm + b1.norm * mean(raindata[t-1,]))
    y.norm[t,] <- ifelse(y.norm[t,] > ut[t,], y.norm[t,], 0)
}
```

#Visualize the replicate

##Skew t

```{r qqplot Skew t}
par(mfrow=c(3,3))
for (i in 1:8){
    plot.ts(raindata[,i], 
            ylim = c(0, 0.1 + max(c(raindata[,i],round(y.st[,i])))))
    lines(y.st[,i],col='grey')
}

    qqplot(c(y.st), c(raindata))
    abline(0,1)

```

##Gaussian
```{r qqplot gaussian}
par(mfrow=c(3,3))
for (i in 1:8){
    plot.ts(raindata[,i], 
            ylim = c(0, 0.1 + max(c(raindata[,i],round(y.norm[,i])))))
    lines(y.norm[,i],col='grey')
}

    qqplot(c(y.norm), c(raindata))
    abline(0,1)

```

##ggplot
```{r skew-t sims}
rain.sims <- rain.long[-(1:8)*28801,]
rain.sims$rain1 <- melt(y.st)$value
rain.sim.plot<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=0.6)+
     geom_line(aes(y = rain1,color='Simulation'),alpha=0.6) + 
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
    ggtitle("(a) Simulation vs Observation" )

rain.sim.plot
```

```{r Gaussian sims}
rain.sims <- rain.long[-(1:8)*28801,]
rain.sims$rain1 <- melt(y.norm)$value
rain.sim.plot.norm<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=0.6)+
     geom_line(aes(y = rain1,color='Simulation'),alpha=0.6) + 
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
    ggtitle("(b) Simulation vs Observation--Guassian Error" )

rain.sim.plot.norm
```


#Function to simulate multiple realizations

##Skew t
```{r multiple simulations}

nsims.app <- function(NT,NS,sigma21,beta1,b0,b1,alpha,nu,niters,true){
    error <- array(0,c(NT,NS,niters))
    y <- array(0,c(NT,NS,niters))
    dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
        }
    }
    B1 <-  sigma21 * matern(dist,beta1,1)
    for(iter in 1: niters){
        for(s in 1 :NS){
            nu <- nu
            bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
            alpha/sqrt(1 + alpha^2)
            omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
            xi<- -omega * bvdelta
            error[,s,iter] <- rst(NT, xi = xi, omega = omega, alpha, nu)
        }

        for(t in 2: NT){
        y[t,,iter] <- B1 %*% true[t - 1,] + 
                            error[t,,iter] * (b0 + b1 * mean(true[t - 1,]))
        y[t,,iter] <- ifelse(y[t,,iter] > ut[t,], y[t,,iter], 0)
        }
    }
    return(y)
}
```

##Gaussian
```{r multiple simulations norm}
nsims.app.norm <- function(NT,NS,sigma21,beta1,b0,b1,niters,true){
    error <- array(0,c(NT,NS,niters))
    y <- array(0,c(NT,NS,niters))
    dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
        }
    }
    B1 <-  sigma21 * matern(dist,beta1,1)
    for(iter in 1: niters){
        for(s in 1 :NS){
            error[,s,iter] <- rnorm(NT)
        }

        for(t in 2: NT){
        y[t,,iter] <- B1 %*% true[t - 1,] + 
                            error[t,,iter] * (b0 + b1 * mean(true[t - 1,]))
        y[t,,iter] <- ifelse(y[t,,iter] > ut[t,], y[t,,iter], 0)
        }
    }
    return(y)
}
```

##Simulation 50 replications
```{r run functions}
set.seed(4)
niters<-50
app.sim <- nsims.app(NT,NS,sigma21.st,beta1.st,b0.st,b1.st,
                  alpha,4,niters,raindata)
app.sim.norm<-nsims.app.norm(NT,NS,sigma21.norm,beta1.norm,b0.norm,b1.norm,
                  niters,raindata)
```

#Calculate the some important quantities related to rainfall

##MRMSE
```{r}
rmse.app<-rmse.app.norm<-c()
for(i in 1:niters){
  rmse.app[i]<-sqrt(mean((app.sim[,,i]-raindata)^2))
  rmse.app.norm[i]<-sqrt(mean((app.sim.norm[,,i]-raindata)^2))
}
mrmse.app<-mean(rmse.app)*100
mrmse.app.norm<-mean(rmse.app.norm)*100
```

##Quantile
```{r}
quan.app<-quan.app.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
   quan.app[j,]<- quantile(app.sim[,,j],1:10000/10000) 
   quan.app.norm[j,]<- quantile(app.sim.norm[,,j],1:10000/10000) 
   quan.app.true<- quantile(raindata,1:10000/10000) 
}
quan.df.app<-data.frame(x=rep(quan.app.true,3),median=c(apply(quan.app,2,sort)[26,],apply(quan.app.norm,2,sort)[26,],quan.app.true),lower=c(apply(quan.app,2,sort)[2,],apply(quan.app.norm,2,sort)[2,],rep(NA,10000)),upper=c(apply(quan.app,2,sort)[49,],apply(quan.app.norm,2,sort)[49,],rep(NA,10000)),Type=c(rep("Skew-t",10000),rep("Gaussian",10000),rep("Truth",10000)))

```

##Simutaneous rain
```{r}
# Mode <- function(x, na.rm = FALSE) {
#   if(na.rm){
#     x = x[!is.na(x)]
#   }
# 
#   ux <- unique(x)
#   return(ux[which.max(tabulate(match(x, ux)))])
# }
simutan<-apply(y.st,1,function(x){sum(x>0)})
#simutan.norm<-apply(y.norm,1,function(x){sum(x>0)})
simutan.true<-apply(raindata,1,function(x){sum(x>0)})



spat.dep <- melt(data.frame(Truth=simutan.true,"Skew-t" = simutan))
```

##Dry spell length
```{r}
cons<-rle(c(raindata))
mattovec<-apply(app.sim,3,c)
mattovec.norm<-apply(app.sim.norm,3,c)
cons2<-apply(mattovec,2,rle)
cons2.norm<-apply(mattovec.norm,2,rle)

quan.cons<-quan.cons.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
   quan.cons[j,]<- quantile(cons2[[j]]$length[cons2[[j]]$values==0],1:10000/10000,type = 6) 
   #quan.cons.norm[j,]<-quantile(cons2.norm[[j]]$length[cons2.norm[[j]]$values==0],1:10000/10000,type = 6)
   quan.cons.true<- quantile(cons$length[cons$values==0],1:10000/10000,type = 6) 
}
quan.cons<-data.frame(x=rep(quan.cons.true,2),median=c(apply(quan.cons,2,sort)[26,],quan.cons.true),lower=c(apply(quan.cons,2,sort)[2,],rep(NA,10000)),upper=c(apply(quan.cons,2,sort)[49,],rep(NA,10000)),Type=c(rep("Skew-t",10000),rep("Truth",10000)))

```

##Wet spell length

```{r}
w.cons<-rle(c(raindata>0))
w.mattovec<-apply(app.sim>0,3,c)
w.mattovec.norm<-apply(app.sim.norm>0,3,c)
w.cons2<-apply(w.mattovec,2,rle)
w.cons2.norm<-apply(w.mattovec.norm,2,rle)

w.quan.cons<-w.quan.cons.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
   w.quan.cons[j,]<- quantile(w.cons2[[j]]$length[w.cons2[[j]]$values==TRUE],1:10000/10000,type = 6) 
   w.quan.cons.norm[j,]<-quantile(w.cons2.norm[[j]]$length[w.cons2.norm[[j]]$values==TRUE],1:10000/10000,type = 6)
   w.quan.cons.true<- quantile(w.cons$length[w.cons$values==TRUE],1:10000/10000,type = 6) 
}
w.quan.cons<-data.frame(x=rep(w.quan.cons.true,2),median=c(apply(w.quan.cons,2,sort)[26,],w.quan.cons.true),lower=c(apply(w.quan.cons,2,sort)[2,],rep(NA,10000)),upper=c(apply(w.quan.cons,2,sort)[49,],rep(NA,10000)),Type=c(rep("Skew-t",10000),rep("Truth",10000)))

```


##Conditional probability
```{r}
condi.prob<-condi.prob.norm<-matrix(0,3,50)
for(i in 1:50){
condi.prob[,i]<-table((app.sim[-1,,i]>0)-(app.sim[-NT,,i]>0))/(NT*NS)
#condi.prob.norm[,i]<-table((app.sim.norm[-1,,i]>0)-(app.sim.norm[-NT,,i]>0))/(NT*NS)
}
condi.prob.real<-table((raindata[-1,]>0)-(raindata[-NT,]>0))/(NT*NS)

wcd<-apply(condi.prob,1,mean)[3]
dcd<-1-apply(condi.prob,1,mean)[3]
wcw<-1-apply(condi.prob,1,mean)[1]
dcw<-apply(condi.prob,1,mean)[1]
round(c(wcw,dcw,wcd,dcd),3)
wcd.r<-apply(condi.prob.real,1,mean)[3]
dcd.r<-1-apply(condi.prob.real,1,mean)[3]
wcw.r<-1-apply(condi.prob.real,1,mean)[1]
dcw.r<-apply(condi.prob.real,1,mean)[1]
round(c(wcw.r,dcw.r,wcd.r,dcd.r),3)
```

#Visualization
##QQplot

```{r}
p1 <- ggplot(data = quan.df.app, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (count/epoch)") + xlab("Sample Quantiles (count/epoch)") +
    ggtitle("QQ Plot of the Observation and Simulation Models") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2")
p1
```

##Simutaneous rain
```{r}

p2 <- ggplot(spat.dep,aes(group=variable,fill=variable)) +
        geom_histogram(aes(value),binwidth=.5,position = position_dodge())+ 
        xlab("Number of simutaneously rainy locations")+
        ylab("Count")+
        ggtitle("Histogram of Simultaneous Rain")
p2

```

##Dry spell length
```{r}
p.cons <- ggplot(data = quan.cons, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (count/epoch)") + xlab("Sample Quantiles (count/epoch)") +
    ggtitle("QQ-plot of the dry spell length") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2")
p.cons

```

```{r}
w.p.cons <- ggplot(data = w.quan.cons, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (count/epoch)") + xlab("Sample Quantiles (count/epoch)") +
    ggtitle("QQ-plot of the wet spell length") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2")
w.p.cons

```

## Distriburion table
```{r distribution table}
tab1<-table(raindata)
tab2<-table(round(y.st))
tab1
tab2

tab1<-table(raindata)
tab2<-table(round(y.norm))
tab1
tab2
```

##Conditional probability

```{r}
par(mfrow=c(1,2))
plot(density(condi.prob[3,]),main='Wet|Dry',xlab="Transition probability",xlim=c(0,0.5),lwd=2,col=2)
lines(density(condi.prob.norm[3,]),col=3,lwd=2,lty=2)
abline(v=condi.prob.real[3],col=1,lwd=2,lty=2)
legend("topright",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2)

plot(density(1-condi.prob[1,]),main='Wet|Wet',xlab="Transition probability",xlim=c(0.5,1),lwd=2,col=2)
lines(density(1-condi.prob.norm[1,]),col=3,lwd=2,lty=2)
abline(v=1-condi.prob.real[1],col=1,lwd=2,lty=2)
legend("topleft",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2)
```

#Save the figures

```{r}
ggsave(filename=here("results","rainsim.pdf"),device='pdf', plot=rain.sim.plot,width = 8, height = 10)
```

```{r}
ggsave(filename=here("results","rainsim.norm.pdf"),device='pdf', plot=rain.sim.plot.norm,width = 8, height = 10)
```

```{r save the figure}
combine_plot<-plot_grid( rain.sim.plot, rain.sim.plot.norm, align='h',rel_widths = 1.2, rel_heights = 1)
ggsave(filename=here("results","sim_compare.pdf"),device='pdf', plot=combine_plot, width = 12, height = 10)
```

```{r}
ggsave(filename=here("results","qqapp.pdf"),device='pdf', plot=p1, width = 6, height = 5)
```

```{r}
ggsave(filename=here("results","spatial.pdf"),device='pdf', plot=p2,width = 6, height = 5)
```

```{r}
ggsave(filename=here("results","length.pdf"),device='pdf', plot=p.cons,width = 6, height = 5)
```

```{r}
ggsave(filename=here("results","wetlength.pdf"),device='pdf', plot=w.p.cons,width = 6, height = 5)
```

```{r,eval=FALSE}
pdf("../results/condidry.pdf",width=10, height=5)
par(mfrow=c(1,2))
plot(density(condi.prob[3,]),main='Wet|Dry',xlab="Transition probability",xlim=c(0,0.5),lwd=2,col=2)
lines(density(condi.prob.norm[3,]),col=3,lwd=2,lty=2)
abline(v=condi.prob.real[3],col=1,lwd=2,lty=2)
legend("topright",c("Truth","Skew-t","Gaussian"),lwd=2,col=c(1,2,3),lty=c(2,1,1))

plot(density(1-condi.prob[1,]),main='Wet|Wet',xlab="Transition probability",xlim=c(0.5,1),lwd=2,col=2)
lines(density(1-condi.prob.norm[1,]),col=3,lwd=2,lty=2)
abline(v=1-condi.prob.real[1],col=1,lwd=2,lty=2)
legend("topleft",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2,lty=c(2,1,1))
dev.off()
```
