---
title: "Inference and Simulation"
author: "Yuxiao Li"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
start_time <- Sys.time()
```


# Overview
This file contains the code for inference and reproducing rainfall data. Most of results shown in the paper can be found here.

In general, we want to

* load the results from estimation procedure
* simulate one replicate of rainfall base on skew-t model and Gaussian model
* visualize the simulations
* simulate multiple replicates of rainfall base on skew-t model and Gaussian model
* calculate the some important quantities related to rainfall
* visualize the results

# Librares and Data

## Libraries
```{r}
library(here)
library(geoR)
library(sn)
#library(cowplot)
```

## load data
```{r load estimate values}
load(here("results", "estimation.cov.VAR1.RData"))
load(here("results", "rain.long.RData"))
raindata <- matrix(rain.long$rain[1:230400],28800,8)
#raindata <- matrix(rain.1min.df$rain[1:115200],14400,8)
NT <- nrow(raindata)
NS <- ncol(raindata)
sigma21.st <- results.VAR1$sigma21
beta1.st <- results.VAR1$beta1
dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
for (i in 1:NS) {
    for (j in 1:NS) {
        dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
    }
}
B1.st <-  sigma21.st * matern(dist,beta1.st,1)
alpha <- results.VAR1$alpha
b0.st <- results.VAR1$b0
b1.st <- results.VAR1$b1
ut <- results.VAR1$ut
```

```{r load estimate values norm}
load(here("results", "estimation.cov.VAR1.norm.RData"))
load(here("results", "rain.long.RData"))
raindata <- matrix(rain.long$rain[1:230400],28800,8)
#raindata <- matrix(rain.1min.df$rain[1:115200],14400,8)
NT <- nrow(raindata)
NS <- ncol(raindata)
sigma21.norm <- results.VAR1.norm$sigma21
beta1.norm <- results.VAR1.norm$beta1
dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
for (i in 1:NS) {
    for (j in 1:NS) {
        dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
    }
}
B1.norm <-  sigma21.norm * matern(dist,beta1.norm,1)
b0.norm <- results.VAR1.norm$b0
b1.norm <- results.VAR1.norm$b1
```


#Simulate one replicate

##Simulate one replicate from skew-t model
```{r simulate 1 error st}
error.st <- matrix(0,NT,NS)
set.seed(4)
for(s in 1 :NS){
    nu <- 4
    bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
        alpha/sqrt(1 + alpha^2)
    omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
    xi<- -omega * bvdelta
    error.st[,s] <- rst(NT, xi = xi, omega = omega, alpha, nu)
}
```

```{r simulate 1 rainfall st}
y.st <- matrix(0,NT,NS)
for(t in 2: NT){
    y.st[t,] <- B1.st %*% raindata[t - 1,] + error.st[t,] * (b0.st + b1.st * mean(raindata[t - 1,]))
    y.st[t,] <- ifelse(y.st[t,] > ut[t,], y.st[t,], 0)
}
```

##Simulate one replicate from Gaussian model

```{r simulate 1 error norm}
error.norm <- matrix(0,NT,NS)
set.seed(4)
for(s in 1 :NS) error.norm[,s] <- rnorm(NT)
```

```{r simulate 1 rainfall norm}
y.norm <- matrix(0,NT,NS)
for(t in 2: NT){
    y.norm[t,] <- B1.norm %*% raindata[t - 1,] + 
        error.norm[t,] * (b0.norm + b1.norm * mean(raindata[t-1,]))
    y.norm[t,] <- ifelse(y.norm[t,] > ut[t,], y.norm[t,], 0)
}
```

#Visualize the replicate

##Skew t

```{r qqplot Skew t, eval=FALSE}
par(mfrow=c(3,3))
for (i in 1:8){
    plot.ts(raindata[,i], 
            ylim = c(0, 0.1 + max(c(raindata[,i],round(y.st[,i])))))
    lines(y.st[,i],col='grey')
}

    qqplot(c(y.st), c(raindata))
    abline(0,1)

```

##Gaussian
```{r qqplot gaussian,eval=FALSE}
par(mfrow=c(3,3))
for (i in 1:8){
    plot.ts(raindata[,i], 
            ylim = c(0, 0.1 + max(c(raindata[,i],round(y.norm[,i])))))
    lines(y.norm[,i],col='grey')
}

    qqplot(c(y.norm), c(raindata))
    abline(0,1)

```

##ggplot
```{r skew-t sims,eval=FALSE}
rain.sims <- rain.long[-(1:8)*28801,]
rain.sims$rain1 <- melt(y.st)$value
rain.sim.plot<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=0.6)+
     geom_line(aes(y = rain1,color='Simulation'),alpha=0.6) + 
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
    ggtitle("(a) Simulation vs Observation" )

rain.sim.plot
```

```{r Gaussian sims,eval=FALSE}
rain.sims <- rain.long[-(1:8)*28801,]
rain.sims$rain1 <- melt(y.norm)$value
rain.sim.plot.norm<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=0.6)+
     geom_line(aes(y = rain1,color='Simulation'),alpha=0.6) + 
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
    ggtitle("(b) Simulation vs Observation--Guassian Error" )

rain.sim.plot.norm
```


#Function to simulate multiple realizations

##Skew t
```{r multiple simulations}

nsims.app <- function(NT,NS,sigma21,beta1,b0,b1,alpha,nu,niters,true){
    error <- array(0,c(NT,NS,niters))
    y <- array(0,c(NT,NS,niters))
    dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
        }
    }
    B1 <-  sigma21 * matern(dist,beta1,1)
    for(iter in 1: niters){
        for(s in 1 :NS){
            nu <- nu
            bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
            alpha/sqrt(1 + alpha^2)
            omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
            xi<- -omega * bvdelta
            error[,s,iter] <- rst(NT, xi = xi, omega = omega, alpha, nu)
        }

        for(t in 2: NT){
        y[t,,iter] <- B1 %*% true[t - 1,] + 
                            error[t,,iter] * (b0 + b1 * mean(true[t - 1,]))
        y[t,,iter] <- ifelse(y[t,,iter] > ut[t,], y[t,,iter], 0)
        }
    }
    return(y)
}
```

##Gaussian
```{r multiple simulations norm}
nsims.app.norm <- function(NT,NS,sigma21,beta1,b0,b1,niters,true){
    error <- array(0,c(NT,NS,niters))
    y <- array(0,c(NT,NS,niters))
    dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
        }
    }
    B1 <-  sigma21 * matern(dist,beta1,1)
    for(iter in 1: niters){
        for(s in 1 :NS){
            error[,s,iter] <- rnorm(NT)
        }

        for(t in 2: NT){
        y[t,,iter] <- B1 %*% true[t - 1,] + 
                            error[t,,iter] * (b0 + b1 * mean(true[t - 1,]))
        y[t,,iter] <- ifelse(y[t,,iter] > ut[t,], y[t,,iter], 0)
        }
    }
    return(y)
}
```

##Simulation 50 replications
```{r run functions}
set.seed(1)
niters<-50
app.sim <- nsims.app(NT,NS,sigma21.st,beta1.st,b0.st,b1.st,
                  alpha,4,niters,raindata)
app.sim.norm <- nsims.app.norm(NT,NS,sigma21.norm,beta1.norm,b0.norm,b1.norm,
                  niters,raindata)
```

#Calculate the some important quantities related to rainfall

##MRMSE
```{r}
rmse.app<-rmse.app.norm<-c()
for(i in 1:niters){
  rmse.app[i] <- sqrt(mean((app.sim[,,i]-raindata)^2))
  rmse.app.norm[i]<-sqrt(mean((app.sim.norm[,,i]-raindata)^2))
}
mrmse.app <- mean(rmse.app) * 100
mrmse.app.norm <- mean(rmse.app.norm) * 100
```

##Quantile
```{r}
quan.app<-quan.app.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
   quan.app[j,]<- quantile(app.sim[,,j],1:10000/10000) 
   quan.app.norm[j,]<- quantile(app.sim.norm[,,j],1:10000/10000) 
   quan.app.true<- quantile(raindata,1:10000/10000) 
}
quan.df.app <- data.frame(x=rep(quan.app.true,3),median=c(apply(quan.app,2,sort)[26,],apply(quan.app.norm,2,sort)[26,],quan.app.true),lower=c(apply(quan.app,2,sort)[2,],apply(quan.app.norm,2,sort)[2,],rep(NA,10000)),upper=c(apply(quan.app,2,sort)[49,],apply(quan.app.norm,2,sort)[49,],rep(NA,10000)),Type=c(rep("Skew-t Simulation",10000),rep("Gaussian Simulation",10000),rep("Observation",10000)))

```

##Simutaneous rain
```{r}
# Mode <- function(x, na.rm = FALSE) {
#   if(na.rm){
#     x = x[!is.na(x)]
#   }
# 
#   ux <- unique(x)
#   return(ux[which.max(tabulate(match(x, ux)))])
# }
simu.mean <- apply(apply(app.sim,c(1,3),function(x){sum(x>0)}),1,mean)
simu.sd <- apply(apply(app.sim,c(1,3),function(x){sum(x>0)}),1,sd)
simu.CI.upper <- simu.mean + 1.96*simu.sd
simu.CI.lower <- simu.mean - 1.96*simu.sd
simu.norm.mean<-apply(apply(app.sim.norm,c(1,3),function(x){sum(x>0)}),1,mean)
simu.norm.sd<-apply(apply(app.sim.norm,c(1,3),function(x){sum(x>0)}),1,sd)
simu.norm.CI.upper <- simu.norm.mean + 1.96*simu.norm.sd
simu.norm.CI.lower <- simu.norm.mean - 1.96*simu.norm.sd

simutan.true<-apply(raindata,1,function(x){sum(x>0)})



spat.dep.true <- data.frame(mean = simutan.true,CI.upper = NA, CI.lower = NA,type= "Observation")
spat.dep.skew <- data.frame(mean = simu.mean,CI.upper = simu.CI.upper, CI.lower = simu.CI.lower,type= "Skew-t Simulation")
spat.dep.norm <- data.frame(mean = simu.norm.mean,CI.upper = simu.norm.CI.upper, CI.lower = simu.norm.CI.lower, type= "Gaussian Simulation")
spat.dep <- rbind(spat.dep.true,spat.dep.skew,spat.dep.norm)

freq_cut <- function(x){
    table(cut(x, breaks=c(-.5,0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5)))
}
```

##Dry spell length
```{r}
cons <- rle(c(raindata))
mattovec <- apply(app.sim,3,c)
mattovec.norm <- apply(app.sim.norm,3,c)
cons2 <- apply(mattovec,2,rle)
cons2.norm <- apply(mattovec.norm,2,rle)

quan.cons <- quan.cons.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
   quan.cons[j,] <- quantile(cons2[[j]]$length[cons2[[j]]$values==0],
                             1:10000/10000,type = 6) 
   quan.cons.true <- quantile(cons$length[cons$values==0],
                              1:10000/10000,type = 6) 
}
quan.cons <- data.frame(x = rep(quan.cons.true,2),
                        median = c(apply(quan.cons,2,sort)[26,],quan.cons.true),
                        lower=c(apply(quan.cons,2,sort)[2,],rep(NA,10000)),
                        upper=c(apply(quan.cons,2,sort)[49,],rep(NA,10000)),
                        Type=c(rep("Skew-t",10000),rep("Truth",10000)))

```

```{r}
plot_multi_histogram <- function(df, feature, label_column) {
    plt <- ggplot(df, aes(x=eval(parse(text=feature)), fill=eval(parse(text=label_column)))) +
    geom_histogram(alpha=0.4, position="identity", aes(y = ..density..),bins=20) +
    labs(x=feature, y = "Density") + scale_x_log10()
    plt + guides(fill=guide_legend(title=label_column))
}
```

```{r}
dry.length<-c()
for (j in 1 :niters){
    dry.length = c(dry.length,cons2[[j]]$length[cons2[[j]]$values==0])
}
dry.length.norm<-c()
for (j in 1 :niters){
    dry.length.norm = c(dry.length,cons2.norm[[j]]$length[cons2.norm[[j]]$values==0])
}

dry.length.true = cons$length[cons$values==0]

a=data.frame(DryLength = (dry.length), Type = "skew-t")
b=data.frame(DryLength = (dry.length.norm), Type = "Normal")
c=data.frame(DryLength = (dry.length.true), Type = "Observation")
abc <- rbind(a,b,c)
plot_multi_histogram(abc,"DryLength","Type")

#abc_boxplot <- ggplot(abc, aes(y=DryLength,group=Type) ) +
#    geom_boxplot(alpha=0.8)+scale_y_log10()
#abc_boxplot
```


##Wet spell length

```{r}
w.cons<-rle(c(raindata>0))
w.mattovec<-apply(app.sim>0,3,c)
w.mattovec.norm<-apply(app.sim.norm>0,3,c)
w.cons2<-apply(w.mattovec,2,rle)
w.cons2.norm<-apply(w.mattovec.norm,2,rle)

w.quan.cons<-w.quan.cons.norm<-matrix(0,niters,10000)
for (j in 1 :niters){
    w.quan.cons[j,] <- quantile(w.cons2[[j]]$length[w.cons2[[j]]$values==TRUE],
                                1:10000/10000,type = 6) 
    w.quan.cons.norm[j,] <- quantile(w.cons2.norm[[j]]$length[w.cons2.norm[[j]]$values==TRUE],
                                     1:10000/10000,type = 6)
    w.quan.cons.true <- quantile(w.cons$length[w.cons$values==TRUE],
                                 1:10000/10000,type = 6) 
}
w.quan.cons <- data.frame(x=rep(w.quan.cons.true,2),
                          median=c(apply(w.quan.cons,2,sort)[26,],w.quan.cons.true),
                          lower=c(apply(w.quan.cons,2,sort)[2,],rep(NA,10000)),
                          upper=c(apply(w.quan.cons,2,sort)[49,],rep(NA,10000)),
                          Type=c(rep("Skew-t",10000),rep("Truth",10000)))

```


##Conditional probability
```{r}
condi.prob <- condi.prob.norm<-matrix(0,3,50)
for(i in 1:50){
condi.prob[,i] <- table((app.sim[-1,,i]>0) - (app.sim[-NT,,i]>0))/(NT*NS)
condi.prob.norm[,i] <- table((app.sim.norm[-1,,i]>0) - (app.sim.norm[-NT,,i]>0))/(NT*NS)
}
condi.prob.real <- table((raindata[-1,]>0)-(raindata[-NT,]>0))/(NT*NS)

wcd <- apply(condi.prob,1,mean)[3]
dcd <- 1-apply(condi.prob,1,mean)[3]
wcw <- 1-apply(condi.prob,1,mean)[1]
dcw <- apply(condi.prob,1,mean)[1]
wcd.sd <- apply(condi.prob,1,sd)[3]
dcd.sd <- apply(condi.prob,1,sd)[3]
wcw.sd <- apply(condi.prob,1,sd)[1]
dcw.sd <- apply(condi.prob,1,sd)[1]
round(c(wcw,dcw,wcd,dcd),3)
round(c(wcw.sd,dcw.sd,wcd.sd,dcd.sd),6)
wcd.n <- apply(condi.prob.norm,1,mean)[3]
dcd.n <- 1-apply(condi.prob.norm,1,mean)[3]
wcw.n <- 1-apply(condi.prob.norm,1,mean)[1]
dcw.n <- apply(condi.prob.norm,1,mean)[1]
wcd.sd.n <- apply(condi.prob.norm,1,sd)[3]
dcd.sd.n <- apply(condi.prob.norm,1,sd)[3]
wcw.sd.n <- apply(condi.prob.norm,1,sd)[1]
dcw.sd.n <- apply(condi.prob.norm,1,sd)[1]
round(c(wcw.n,dcw.n,wcd.n,dcd.n),3)
round(c(wcw.sd.n,dcw.sd.n,wcd.sd.n,dcd.sd.n),6)
wcd.r <- condi.prob.real[3]
dcd.r <- 1-condi.prob.real[3]
wcw.r <- 1-condi.prob.real[1]
dcw.r <- condi.prob.real[1]
round(c(wcw.r,dcw.r,wcd.r,dcd.r),3)
```

#Visualization
##QQplot

```{r}
p1 <- ggplot(data = quan.df.app, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (mm/hr)") + xlab("Sample Quantiles (mm/hr)") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2") + guides(colour=guide_legend(title="Data Type"),fill=guide_legend(title="Data Type"))
p1
```

##Simutaneous rain
```{r}

p2 <- ggplot(spat.dep,aes(x = mean, fill=type)) +
        geom_histogram(alpha=0.8, position="dodge2", aes(y = ..density..),bins=9) +
        scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8)) + scale_y_sqrt()+
        xlab("Number of simutaneously rainy locations") +
        ylab("Frequency") + guides(fill=guide_legend(title="Data Type"))

p2


```

##Dry spell length
```{r}
p.cons <- ggplot(data = quan.cons, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (count/epoch)") + xlab("Sample Quantiles (count/epoch)") +
    ggtitle("QQ-plot of the dry spell length") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2")
p.cons
```



```{r}
w.p.cons <- ggplot(data = w.quan.cons, aes(x=x, y=median, colour=Type)) + 
    geom_line(size=0.8) + ylab("Empirical Quantiles (count/epoch)") + xlab("Sample Quantiles (count/epoch)") +
    ggtitle("QQ-plot of the wet spell length") +
    scale_y_continuous()+scale_x_continuous()+
    geom_ribbon(aes(ymin=lower, ymax=upper,fill=Type), linetype=2, alpha=0.3) + 
    scale_color_brewer(palette = "Dark2") +    
    scale_fill_brewer(palette = "Dark2")
```

## Distriburion table
```{r distribution table}
tab1<-table(raindata)
tab2<-table(round(y.st))
tab1
tab2

tab1<-table(raindata)
tab2<-table(round(y.norm))
tab1
tab2
```

##Conditional probability

```{r,eval = FALSE}
par(mfrow=c(1,2))
plot(density(condi.prob[3,]),main='Wet|Dry',
     xlab="Transition probability",xlim=c(0,0.5),lwd=2,col=2)
lines(density(condi.prob.norm[3,]),col=3,lwd=2,lty=2)
abline(v=condi.prob.real[3],col=1,lwd=2,lty=2)
legend("topright",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2)

plot(density(1-condi.prob[1,]),main='Wet|Wet',
     xlab="Transition probability",xlim=c(0.5,1),lwd=2,col=2)
lines(density(1-condi.prob.norm[1,]),col=3,lwd=2,lty=2)
abline(v=1-condi.prob.real[1],col=1,lwd=2,lty=2)
legend("topleft",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2)
```

#Save the figures

```{r}
ggsave(filename=here("results","rainsim.pdf"),device='pdf', plot=rain.sim.plot,width = 8, height = 10)
```

```{r}
ggsave(filename=here("results","rainsim.norm.pdf"),device='pdf', plot=rain.sim.plot.norm,width = 8, height = 10)
```

```{r save the figure}
combine_plot<-plot_grid( rain.sim.plot, rain.sim.plot.norm, align='h',rel_widths = 1.2, rel_heights = 1)
ggsave(filename=here("results","sim_compare.pdf"),device='pdf', plot=combine_plot, width = 12, height = 10)
```

```{r}
ggsave(filename=here("results","qqapp.pdf"),device='pdf', plot=p1, width = 8, height = 5)
```

```{r}
ggsave(filename=here("results","spatial.pdf"),device='pdf', plot=p2,width = 8, height = 5)
```

```{r}
ggsave(filename=here("results","length.pdf"),device='pdf', plot=p.cons,width = 6, height = 5)
```

```{r}
ggsave(filename=here("results","wetlength.pdf"),device='pdf', plot=w.p.cons,width = 6, height = 5)
```

```{r,eval=FALSE}
pdf("../results/condidry.pdf",width=10, height=5)
par(mfrow=c(1,2))
plot(density(condi.prob[3,]),main='Wet|Dry',xlab="Transition probability",xlim=c(0,0.5),lwd=2,col=2)
lines(density(condi.prob.norm[3,]),col=3,lwd=2,lty=2)
abline(v=condi.prob.real[3],col=1,lwd=2,lty=2)
legend("topright",c("Truth","Skew-t","Gaussian"),lwd=2,col=c(1,2,3),lty=c(2,1,1))

plot(density(1-condi.prob[1,]),main='Wet|Wet',xlab="Transition probability",xlim=c(0.5,1),lwd=2,col=2)
lines(density(1-condi.prob.norm[1,]),col=3,lwd=2,lty=2)
abline(v=1-condi.prob.real[1],col=1,lwd=2,lty=2)
legend("topleft",c("Truth","Skew-t","Gaussian"),col=c(1,2,3),lwd=2,lty=c(2,1,1))
dev.off()
```

## Running time
```{r, echo = FALSE}
end_time <- Sys.time()

runtime_05 <- end_time-start_time

runtime_05
```