---
title: "Prediction"
author: "Yuxiao Li"
output: 
  github_document:
    toc: true
  html_document:
    toc: true
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
start_time <- Sys.time()
```

# Overview
This file contains the code for prediction.

In general, we want to
* split the dataset to training and testing set
* transfer the dataset to python to run the LSTM model
* load the predicted value from LSTM model
* calculate the predicted value from our model
* compare the MSE and other important properties.

# Librares and Data

## Libraries

```{r load_libs,message=FALSE}
library(here)
library(sn)
library(geoR)
library(reshape2)
library(nloptr)
```

## Data

Different resolution can be used for the analysis, the default setting is to use the rawdata, but we can also use the aggregated data in 1-minute, 15-minute, and hourly resolution.

```{r load_data}

load(here("results", "rain.long.RData"))
raindata <- matrix(rain.long$rain[1:230400],28800,8)
sep <- 1
#load(here("results", "rain.1min.df.RData"))
#raindata <- matrix(rain.1min.df$rain[1:115200],14400,8)
#sep <- 2
#load(here("results", "rain.15min.df.RData"))
#raindata <- matrix(rain.15min.df$rain[1:7680],960,8)
#sep <- 30
##load(here("results", "rain.hour.df.RData"))
#raindata <- matrix(rain.hour.df$rain[1:1920],240,8)
#sep <- 120
```

## Save the data as csv to be used in Python
```{r}
write.csv(rain.long,file=here("results", "rain.raw.df.csv"))
```

#Required Functions
## Fit using matern covariance function and skew-t error
```{r likelihood function skew t}
varlik.sp.var1.st <- function(params, data, coord, ut, nu){
    NS <- ncol(data)
    NT <- nrow(data)
    sigma21 <- exp(params[1])/(1+exp(params[1]))  # The variance parameter B1
    beta1 <- exp(params[2])    # The spatial range parameter B1
    b0 <- exp(params[3])
    b1 <- exp(params[4])
    alpha <- exp(params[5]) #Skewness parameter
    nu <- nu ## Degree of freedom
    # Compute the variograms under the model
    dist <- matrix(0, NS, NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- 
                sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)
        }
    }
    # The covariance matrix under model with nugget effect
    B1 <- sigma21 * matern(dist,beta1,1)
    #if(eigen(B1)$values[1] >= 0.99) {
     #   warning("The process is not stationary, return -1e16")
      #  return(1e16)
       # }
    sigt_plus <- b0 + b1 * apply(data,1,mean)
    
    ##Construct the likelihood function
    sum.lik <- 0
    data.lag1 <- data[-1,]
    for (s in 1:NS){
        index.posi <- which(data.lag1[,s] > 0)
        index.zero <- which(data.lag1[,s] == 0)
        x <- (data.lag1[index.posi,s] - data[index.posi,] %*% B1[s,])/
            sigt_plus[index.posi]
        u <- (ut[index.zero + 1, s] - data[index.zero,] %*% B1[s,])/
            sigt_plus[index.zero]
        bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
         alpha/sqrt(1 + alpha^2)
        omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
        xi<- -omega * bvdelta
        part1 <- dst(x, xi, omega, alpha, nu, log=TRUE) #First term
        part2 <- log(sigt_plus[index.posi]) #Second term
        part3 <- log(pst(u, xi, omega, alpha, nu)) #Third term
        sum.lik <- sum.lik + sum(part1) - sum(part2) + sum(part3)
    }
        #message(c(sigma21,beta1,alpha,b0,b1))
    return(-sum.lik)
}
```

## Split the data
```{r}
N.train = 2*60*24*9
N.test = 2*60*24
raindata.train <- raindata[1:N.train,]
raindata.test <- raindata[(N.train+1):(N.train + N.test),]
```

##Fit the training data
```{r estimate skew t}
NT <- nrow(raindata)
NS <- ncol(raindata)

init <- c()
#init[1] <- 0.1/(1/NS-0.1) #variance
init[1] <- 0.1
init[2] <- 1 #spatial range
init[3] <- .5
init[4] <- .5
init[5] <- 1 #positive skewness alpha
init <- log(init)
#cov.optim.rslt.st <- optim(par=init,fn = varlik.sp.var1.st,
#                    data = raindata, coord = coord, ut = ut, nu = 4, 
#                    hessian = TRUE)
#cov.optim.rslt.st.1 <- bobyqa(x0=init,fn = varlik.sp.var1.st,
#                    data = raindata, coord = coord, ut = ut, nu = 4)
cov.optim.rslt.st.1 <- neldermead(x0=init,fn = varlik.sp.var1.st,
                    data = raindata.train, coord = coord, ut = ut[1:N.train,], nu = 4)
#cov.optim.rslt.st.1 <- lbfgs(x0=init,fn = varlik.sp.var1.st,
#                    data = raindata, coord = coord, ut = ut, nu = 4)
```


##Return the estimated values
```{r est results spatial skew t}
cov.optim.rslt <- cov.optim.rslt.st.1

# The covariance matrix under model with nugget effect
sigma21 <- exp(cov.optim.rslt$par[1])/(1+exp(cov.optim.rslt$par[1]))
#cov.sigma21 <- exp(cov.optim.rslt$par[1])
beta1 <- exp(cov.optim.rslt$par[2])
b0 <- exp(cov.optim.rslt$par[3])
b1 <- exp(cov.optim.rslt$par[4])
alpha <- exp(cov.optim.rslt$par[5])
```

#Prediction

##Conditional simulation function
```{r multiple simulations}

nsims.app <- function(NT,NS,sigma21,beta1,b0,b1,alpha,nu,niters,true,ut){
    error <- array(0,c(NT,NS,niters))
    y <- array(0,c(NT,NS,niters))
    dist <- matrix(data = 0, nrow = NS, ncol = NS) # The distances between locations
    for (i in 1:NS) {
        for (j in 1:NS) {
            dist[i,j] <- sqrt((coord[i,2]-coord[j,2])^2+(coord[i,3]-coord[j,3])^2)   
        }
    }
    B1 <-  sigma21 * matern(dist,beta1,1)
    for(iter in 1: niters){
        for(s in 1 :NS){
            nu <- nu
            bvdelta <- sqrt(nu) * gamma(1/2 * (nu - 1))/sqrt(pi)/gamma(1/2 * nu) * 
            alpha/sqrt(1 + alpha^2)
            omega <- 1/sqrt((nu/(nu-2) - bvdelta^2))
            xi<- -omega * bvdelta
            error[,s,iter] <- rst(NT, xi = xi, omega = omega, alpha, nu)
        }

        for(t in 2: NT){
        y[t,,iter] <- B1 %*% true[t - 1,] + 
                            error[t,,iter] * (b0 + b1 * mean(true[t - 1,]))
        y[t,,iter] <- ifelse(y[t,,iter] > ut[t,], y[t,,iter], 0)
        }
    }
    return(y)
}
```

##Predition from our model
```{r}
niters<-50
app.sim <- nsims.app(N.test,NS,sigma21,beta1,b0,b1,
                  alpha,4,niters,raindata.test,ut=ut[(N.train+1):(N.train + N.test),])
```

```{r}
est.LSTM <- read.csv(here("results", "rainest.csv"),header = FALSE)
```

##Visualize the results
```{r}
plot.ts((apply(app.sim,c(1,2),mean)))
plot.ts(raindata.test)
plot.ts(est.LSTM)
```

```{r}
rain.raw <- read.csv(here("data","RaindataPerepoch.csv"))
rain.long <- melt(rain.raw[(N.train+1):(N.train + N.test),1:9], id=c("date...time"))
rain.long <- data.frame(rain.long)

names(rain.long)<-c("dt","sites","rain")
rain.long$dt <- as.POSIXct(strptime(rain.long$dt,format="%d.%m.%Y  %H:%M:%S"))
rain.long$Date <- as.Date(rain.long$dt)
rain.long$Time <- format((rain.long$dt) ,format = "%H:%M:%S")
rain.sims <- rain.long[1:(2877*8),]
rain.sims$rain <- melt(raindata.test[4:2880])$value
rain.sims$LSTM <- melt(est.LSTM)$value
rain.sims$PG_mean <- melt(apply(app.sim[3:2879,,],c(1,2),mean))$value
rain.sims$PG_up <- melt(apply(app.sim[3:2879,,],c(1,2),quantile,0.975))$value
rain.sims$PG_low <- melt(apply(app.sim[3:2879,,],c(1,2),quantile,0.025))$value
rain.sim.plot<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=0.8)+
     geom_line(aes(y = LSTM,color='LSTM Prediction'),alpha=0.8) + 
        geom_line(aes(y = PG_mean,color='Skew-t Prediction'),alpha=0.8) + 
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
        guides(colour=guide_legend(title="Data Type",reverse =TRUE))

rain.sim.plot
```

```{r}
rain.raw <- read.csv(here("data","RaindataPerepoch.csv"))
rain.long <- melt(rain.raw[(N.train+800):(N.train + N.test-1804),1:9], id=c("date...time"))
rain.long <- data.frame(rain.long)

names(rain.long)<-c("dt","sites","rain")
rain.long$dt <- as.POSIXct(strptime(rain.long$dt,format="%d.%m.%Y  %H:%M:%S"))
rain.long$Date <- as.Date(rain.long$dt)
rain.long$Time <- format((rain.long$dt) ,format = "%H:%M:%S")
rain.sims <- rain.long[1:((2877-2600)*8),]
rain.sims$rain <- melt(raindata.test[(800+4):(2880-1800)])$value
rain.sims$LSTM <- melt(est.LSTM[801:(2877-1800),])$value
rain.sims$PG_mean <- melt(apply(app.sim[(3+800):(2879-1800),,],c(1,2),mean))$value
rain.sims$PG_up <- melt(apply(app.sim[(3+800):(2879-1800),,],c(1,2),quantile,0.975))$value
rain.sims$PG_low <- melt(apply(app.sim[(3+800):(2879-1800),,],c(1,2),quantile,0.025))$value

rain.sim.plot1<-ggplot(rain.sims,aes(x=dt,group=sites))+
    geom_line(aes(y = rain,color='Observation'),alpha=1)+
     geom_line(aes(y = PG_mean,color='Skew-t Prediction'),alpha=1) + 
        geom_line(aes(y = LSTM,color='LSTM Prediction'),alpha=1) + 
        geom_ribbon(aes(ymin=PG_low, ymax=PG_up,fill = "Skew-t Prediction"), alpha=0.2)+
        facet_grid(sites ~ .) + xlab("Data and Time") + ylab("Amount of Rain")+
         guides(colour=guide_legend(title="Data Type"),
                fill=guide_legend(title="Prediction Band")) + 
                scale_fill_manual(values=c("blue"), name="fill")

rain.sim.plot1
```


##Calculate MSE
```{r}
mse.pg = mean((apply(app.sim,c(1,2),mean)[-2880,]-raindata.test[-1,])^2)
mse.pg.single <- c()
for(i in 1:50){
    mse.pg.single[i] = mean((app.sim[-2880,,i]-raindata.test[-1,])^2)
}


mse.LSTM = sum((est.LSTM - raindata.test[-c(1,2,3),])^2)/2877/8
```

```{r}
ggsave(filename=here("results","pred1.pdf"),device='pdf', plot=rain.sim.plot,width = 8, height = 6)
```

```{r}
ggsave(filename=here("results","pred2.pdf"),device='pdf', plot=rain.sim.plot1,width = 8, height = 6)
```

